---
title: "Usage of the yellow taxi in January 2019"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction

In New York City the transportation is divided between public transportation such as metropolitan or buses and private transportation where we can find taxis and ride-hailing companies.
If we focus only on taxis in NYC we can find two main group:  
a) the (iconic) **yellow cab** that are concentrated in the borough of Manhattan, but can be hailed anywhere throughout the five boroughs of New York City and may be hailed with a raised hand or by standing at a taxi stand;  
b) the (apple) **green cab** that can be hailed only in the outer boroughs (except at the airports) and in the northern part of Manhattan, specifically above 96th street on the east side and above 110th street on the west side.  
  
The scope of this project is to analyse the most frequent route made by the yellow taxy and which were the most common zone where taxi picks-up or drops-off the people during the month of January 2019.  
The data used are avaiable for free from the NYC Govermnet website at the following link:   https://www1.nyc.gov/site/tlc/about/tlc-trip-record-data.page

## Getting ready  

The first step is to install the libraries useful for the analysis and the visualization.  
For doing this we can use the comand *install.packages('name_of_the_library')*.  
In particular the library *sf* is going to be used for creating the map of New York City while the *networkD3*, a Javascript based library, is going to used for the Sankey Diagram.

```{r}
install.packages('tidyverse')
install.packages('dplyr')
install.packages('sf')
install.packages('networkD3')
install.packages('ggdark')
```

Once installed the libraries we load them using the command *library('name_of_the_library')*.  

```{r}
library(tidyverse)
library(dplyr)
library(sf)
library(networkD3)
library(ggdark)
library(htmltools)
```

## Data preprocessing

Using the command *read_csv*, we load the data about all the trips made by the yellow taxi during the month of January 2019 while with *read_sf* we load the geographical shape file of New York City that will use to create the map of the city.

```{r}
yellow_df <- read_csv("../data/yellow_tripdata_2019-01.csv")
map_nyc <- read_sf("../data/taxi_zones.shp")
```
Let's display the first rows of the data frame yellow_df.  
The comand *head(data_frame)* is useful to start investigating of which column the data frame is composed and which data they contain.

```{r}
head(yellow_df)
```

By this first simple investication we see that the data frame contains two columns that we can use for answering which are the most common zone where taxi pick-up or drop-off. These two coloumn are:
- PULocationID, location where the meter was engaged;  
- DOLocationID, location where  the meter was disengaged;  
  
Let's now displaying the first rows of map_nyc.
```{r}
head(map_nyc)
```
As we can see, also here there is the information about the location (LocationID) what we need to do is to create two columns called 'PULocationID' and 'DOLocatioID'. Why this? So we can later join the data frame about the trips made with this one in this way we can star analyse what are the most common zones and which are the most frequent route.  

```{r}
map_nyc$PULocationID <- map_nyc$LocationID
map_nyc$DOLocationID <- map_nyc$LocationID
```

The dataframe about the yellow taxi taken in January contains about two million rows, for the purpose of the visualization project I decided to take two samples one for each visualization.  
The Sankey diagram is going to be created on a sample of about 30% of the original data set, while the maps are going to be create using a sample of about 0.1% of the original size.    
For both the two new data frames *yellow_df_sample_sankey* and *yellow_df_sample_taxy* we are going to create a subset keeping only the two columns PULocationID, DOLocationID and dropping the rows with NA value.

```{r}
yellow_df_sample_sankey <- yellow_df[sample(nrow(yellow_df), 743219), ] %>% 
  subset(select = c(PULocationID,DOLocationID)) %>%
  drop_na()

yellow_df_sample_taxi <- yellow_df[sample(nrow(yellow_df), 3219), ] %>% 
  subset(select = c(PULocationID,DOLocationID)) %>%
  drop_na()
```

Since want we want to visualize are the zones where most NYC citizen are pick-up or drop-off by a taxi it is better to created two new data frames one for each action. Each data frame will contain only the columns useful to the creation of the map, this can be achive by creating a subset.
  
The two new data frames are made using a left join between the *yellow_df_sample* and *map_nyc*. From this join we only keep the columns: PULocationID (or DOLocationID),Shape_Leng, zone, borough, geometry and we will dropp the rows where NA value is shown.

```{r}
yellow_nyc_PUL <- left_join(yellow_df_sample_taxi,map_nyc,"PULocationID") %>%
  subset(select = c(PULocationID,Shape_Leng,zone,borough, geometry)) %>%
  drop_na()

yellow_nyc_DOL <- left_join(yellow_df_sample_taxi,map_nyc,"DOLocationID") %>%
  subset(select = c(DOLocationID,Shape_Leng,zone,borough, geometry)) %>%
  drop_na()
```

## Data Visualization 

#### Map of the most pick-up location 
The first visualization we are goint to do is for answering the question "Which are the zone where taxi pick-up most people?".  
To answer this question we have to use the data frame *yellow_nyc_PUL*. On this data frame we are going to apply a *group_by* function to the column PULocationID. Once grouped we are going to count how many PULocationID are in the data fram and save this result in a new column called count. The result of this part of the code is saved in a new variable called *yellow_PUL_sample_taxy_map*. 
Once this data frame is created we can join it with the data frame *map_nyc* by the column PULocationID and keeping only a subset of it. The result is going to be saved in a new variable called *yellow_taxi_PUL_map*.  
The map we are going to create should visualize what are the most common zone where taxi picks-up people. For doing this we can use the nyc_map and fill the zone with the count value.

```{r}
yellow_PUL_sample_taxi_map <- yellow_nyc_PUL %>% group_by(PULocationID) %>%
  summarize(count=n())

yellow_taxi_PUL_map <- left_join(yellow_PUL_sample_taxi_map,map_nyc, 'PULocationID') %>%
  subset(select = c(PULocationID, count, geometry))

PUL_map <- ggplot() +
  geom_sf (data = map_nyc, size = 0.03) +
  geom_sf (data = yellow_taxi_PUL_map, mapping = aes(fill = count, geometry = geometry), size = 0.05) +
  scale_fill_distiller(palette = "Spectral", breaks=c(0,25,50,75,100,125,150)) +
  guides(fill = guide_colourbar(barheight = unit( 3 , "in" ), ticks.colour = "black", ticks.linewidth = 1))+
  ggtitle('Which are the zones where taxy picks-up most people?') +
  dark_mode()

PUL_map
```
What is possible to see in the maps is that as expected the yellow taxi operates more intensely inside the Manhattan area in particular around the Upper East Side South, Upper East Side North, the Midtown East and the Midtown Center. Outside of the Manhattan area one zone where yellow taxi picks-up most people is the airport JFK (bottom right).


#### Map of the most drop-off location 

For answering the question about which are the most common zones where taxi drop-off people, we are going to work as before but this time we will consider the column DOLocationID.

```{r}
yellow_DOL_sample_taxi_map <- yellow_nyc_DOL %>% group_by(DOLocationID) %>%
  summarize(count=n())

yellow_taxi_DOL_map <- left_join(yellow_DOL_sample_taxi_map,map_nyc, 'DOLocationID') %>%
  subset(select = c(DOLocationID, count, geometry))

DOL_map <- ggplot() +
  geom_sf (data = map_nyc, size = 0.03) +
  geom_sf (data = yellow_taxi_DOL_map, mapping = aes(fill = count, geometry = geometry), size = 0.05) + 
  scale_fill_distiller(palette = "Spectral", breaks=c(0,25,50,75,100,125,150)) +
  guides(fill = guide_colourbar(barheight = unit( 3 , "in" ), ticks.colour = "black", ticks.linewidth = 1))+
  ggtitle('Which are the zones taxy drops-off most people?') +
  dark_mode()

DOL_map 
```
What we can see from the map above, is that as before inside the Manhanttan zone the yellow taxi drop-off is intense. But we can see that also outside, excluding the JFK airport, yellow taxi are dropping-off people. An assumption of this could be that people that work inside the Manhattan zone lived outside.  




### Sankey diagram 

Script for creating the Sankey Diagram
```{r}
names <- c(
  `1` <- "Newark Airport",
  `2` <- "Jamaica Bay",
  `3` <- "Allerton/Pelham Gardens",
  `4` <- "Arden City",
  `5` <- "Arden Heights",
  `6` <- "Arrochar/Fort Wadsworth",
  `7` <- "Astoria",
  `8` <- "Astoria Park",
  `9` <- "Auburndale",
  `10` <- "Baisley Park",
  `11` <- "Bath Beach",
  `12` <- "Battery Park",
  `13` <- "Battery Park City",
  `14` <- "Bay Ridge",
  `15` <- "Bay Terrace/Fort Totten",
  `16` <- "Bayside",
  `17` <- "Bedford",
  `18` <- "Bedford Park",
  `19` <- "Bellerose",
  `20` <- "Belmont",
  `21` <- "Bensonhurst East",
  `22` <- "Bensonhurst West",
  `23` <- "Bloomfield/Emerson Hill",
  `24` <- "Bloomingdale",
  `25` <- "Boerum Hill",
  `26` <- "Borough Park",
  `27` <- "Breezy Point/Fort Tilden/Riis Beach",
  `28` <- "Briarwood/Jamaica Hills",
  `29` <- "Brighton Beach",
  `30` <- "Broad Channel",
  `31` <- "Bronx Park",
  `32` <- "Bronxdale",
  `33` <- "Brooklyn Heights",
  `34` <- "Brooklyn Navy Yard",
  `35` <- "Brownsville",
  `36` <- "Bushwick North",
  `37` <- "Bushwick South",
  `38` <- "Cambria Heights",
  `39` <- "Canarsie",
  `40` <- "Carroll Gardens",
  `41` <- "Central Harlem",
  `42` <- "Central Harlem North",
  `43` <- "Central Park",
  `44` <- "Charleston/Tottenville",
  `45` <- "Chinatown",
  `46` <- "City Island",
  `47` <- "Claremont/Bathgate",
  `48` <- "Clinton East",
  `49` <- "Clinton Hill",
  `50` <- "Clinton Wes",
  `51` <- "Co-Op City",
  `52` <- "Cobble Hill",
  `53` <- "College Point",
  `54` <- "Columbia Street",
  `55` <- "Coney Island",
  `56` <- "Corona North",
  `57` <- "Corona South",
  `58` <- "Country Clubt",
  `59` <- "Crotona Park",
  `60` <- "Crotona Park Eas",
  `61` <- "Crown Heights North",
  `62` <- "Crown Heights South",
  `63` <- "Cypress Hills",
  `64` <- "Douglaston",
  `65` <- "Downtown Brooklyn/MetroTech",
  `66` <- "DUMBO/Vinegar Hill",
  `67` <- "Dyker Heights",
  `68` <- "East Chelsea",
  `69` <- "East Concourse/Concourse Village",
  `70` <- "East Elmhurst",
  `71` <- "East Flatbush/Farragut",
  `72` <- "East Flatbush/Remsen Village",
  `73` <- "East Flushing",
  `74` <- "East Harlem North",
  `75` <- "East Harlem South",
  `76` <- "East New York",
  `77` <- "East New York/Pennsylvania Avenue",
  `78` <- "East Tremont",
  `79` <- "East Village",
  `80` <- "East Williamsburg",
  `81` <- "Eastchester",
  `82` <- "Elmhurst",
  `83` <- "Elmhurst/Maspeth",
  `84` <- "Eltingville/Annadale/Prince's Bay",
  `85` <- "Erasmus",
  `86` <- "Far Rockaway",
  `87` <- "Financial District North",
  `88` <- "Financial District South",
  `89` <- "Flatbush/Ditmas Park",
  `90` <- "Flatiron",
  `91` <- "Flatlands",
  `92` <- "Flushing",
  `93` <- "Flushing Meadows-Corona Park",
  `94` <- "Fordham South",
  `95` <- "Forest Hills",
  `96` <- "Forest Park/Highland Park",  
  `97` <- "Fort Greene",
  `98` <- "Fresh Meadows",
  `99` <- "Freshkills Park",
  `100` <- "Garment District",
  `101` <- "Glen Oaks",
  `102` <- "Glendale",
  `103` <- "Governor's Island",
  `104` <- "Ellis Island",
  `105` <- "Liberty Island",
  `106` <- "Gowanus",
  `107` <- "Gramercy",
  `108` <- "Gravesend",
  `109` <- "Great Kills",
  `110` <- "Great Kills Park",
  `111` <- "Green-Wood Cemetery",
  `112` <- "Greenpoint",
  `113` <- "Greenwich Village North",
  `114` <- "Greenwich Village South",
  `115` <- "Grymes Hill/Clifton",
  `116` <- "Hamilton Heights",
  `117` <- "Hammels/Arverne",
  `118` <- "Heartland Village/Todt Hill",
  `119` <- "Highbridge",
  `120` <- "Highbridge Park",
  `121` <- "Hillcrest/Pomonok",
  `122` <- "Hollis",
  `123` <- "Homecrest",
  `124` <- "Howard Beach",
  `125` <- "Hudson Sq",
  `126` <- "Hunts Point",
  `127` <- "Inwood",
  `128` <- "Inwood Hill Park",
  `129` <- "Jackson Heights",
  `130` <- "Jamaica",
  `131` <- "Jamaica Estates",
  `132` <- "JFK Airport",
  `133` <- "Kensington",
  `134` <- "Kew Gardens",
  `135` <- "Kew Gardens Hills",
  `136` <- "Kingsbridge Heights",
  `137` <- "Kips Bay",
  `138` <- "LaGuardia Airport",
  `139` <- "Laurelton",
  `140` <- "Lenox Hill East",
  `141` <- "Lenox Hill West",
  `142` <- "Lincoln Square East",
  `143` <- "Lincoln Square West",
  `144` <- "Little Italy/NoLiTa",
  `145` <- "Long Island City/Hunters Point",
  `146` <- "Long Island City/Queens Plaza",
  `147` <- "Longwood",
  `148` <- "Lower East Side",
  `149` <- "Madison",
  `150` <- "Manhattan Beach",
  `151` <- "Manhattan Valley",
  `152` <- "Manhattanville",
  `153` <- "Marble Hill",
  `154` <- "Marine Park/Floyd Bennett Field",
  `155` <- "Marine Park/Mill Basin",
  `156` <- "Mariners Harbor",
  `157` <- "Maspeth",
  `158` <- "Meatpacking/West Village West",
  `159` <- "Melrose South",
  `160` <- "Middle Villagee",
  `161` <- "Midtown Center",
  `162` <- "Midtown East",
  `163` <- "Midtown North",
  `164` <- "Midtown South",
  `165` <- "Midwood",
  `166` <- "Morningside Heights",
  `167` <- "Morrisania/Melrose",
  `168` <- "Mott Haven/Port Morris",
  `169` <- "Mount Hope",
  `170` <- "Murray Hill",
  `171` <- "Murray Hill-Queens",
  `172` <- "New Dorp/Midland Beach",
  `173` <- "North Corona",
  `174` <- "Norwood",
  `175` <- "Oakland Gardens",
  `176` <- "Oakwood",
  `177` <- "Ocean Hill",
  `178` <- "Ocean Parkway South",
  `179` <- "Old Astoria",
  `180` <- "Ozone Park",
  `181` <- "Park Slop",
  `182` <- "Parkchester",
  `183` <- "Pelham Bay",
  `184` <- "Pelham Bay Park",
  `185` <- "Pelham Parkway",
  `186` <- "Penn Station/Madison Sq West",
  `187` <- "Port Richmond",
  `188` <- "Prospect-Lefferts Gardens",
  `189` <- "Prospect Heights",
  `190` <- "Prospect Park",
  `191` <- "Queens Village",
  `192` <- "Queensboro Hill",
  `193` <- "Queensbridge/Ravenswood",
  `194` <- "Randalls Island",
  `195` <- "Red Hook",
  `196` <- "Rego Park",
  `197` <- "Richmond Hill",
  `198` <- "Ridgewood",
  `199` <- "Rikers Island",
  `200` <- "Riverdale/North Riverdale/Fieldston",
  `201` <- "Rockaway Park",
  `202` <- "Roosevelt Island",
  `203` <- "Rosedale",
  `204` <- "Rossville/Woodrow",
  `205` <- "Saint Albans",
  `206` <- "Saint George/New Brighton",
  `207` <- "Saint Michaels Cemetery/Woodside",
  `208` <- "Schuylerville/Edgewater Park",
  `209` <- "Seaport",
  `210` <- "Sheepshead Bay",
  `211` <- "SoHo",
  `212` <- "Soundview/Bruckner",
  `213` <- "Soundview/Castle Hill",
  `214` <- "South Beach/Dongan Hills",
  `215` <- "South Jamaica",
  `216` <- "South Ozone Park",
  `217` <- "South Williamsburg",
  `218` <- "Springfield Gardens North",
  `219` <- "Springfield Gardens South",
  `220` <- "Spuyten Duyvil/Kingsbridge",
  `221` <- "Stapleton",
  `222` <- "Starrett City",
  `223` <- "Steinway",
  `224` <- "Stuy Town/Peter Cooper Village",
  `225` <- "Stuyvesant Heights",
  `226` <- "Sunnyside",
  `227` <- "Sunset Park East",
  `228` <- "Sunset Park West",
  `229` <- "Sutton Place/Turtle Bay North",
  `230` <- "Times Sq/Theatre District",
  `231` <- "TriBeCa/Civic Center",
  `232` <- "Two Bridges/Seward Park",
  `233` <- "UN/Turtle Bay South",
  `234` <- "Union Sq",
  `235` <- "University Heights/Morris Heights",
  `236` <- "Upper East Side North",
  `237` <- "Upper East Side South",
  `238` <- "Upper West Side North",
  `239` <- "Upper West Side South",
  `240` <- "Van Cortlandt Park",
  `241` <- "Van Cortlandt Village",
  `242` <- "Van Nest/Morris Park",
  `243` <- "Washington Heights North",
  `244` <- "Washington Heights South",
  `245` <- "West Brighton",
  `246` <- "West Chelsea/Hudson Yards",
  `247` <- "West Concourse",
  `248` <- "West Farms/Bronx River",
  `249` <- "West Village",
  `250` <- "Westchester Village/Unionport",
  `251` <- "Westerleigh",
  `252` <- "Whitestone",
  `253` <- "Willets Point",
  `254` <- "Williamsbridge/Olinville",
  `255` <- "Williamsburg (North Side)",
  `256` <- "Williamsburg (South Side)",
  `257` <- "Windsor Terrace",
  `258` <- "Woodhaven",
  `259` <- "Woodlawn",
  `260` <- "Wakefield",
  `261` <- "Woodside",
  `262` <- "World Trade Center",
  `263` <- "Yorkville East",
  `264` <- "Yorkville West"
)
``` 

```{r}
final_table <- yellow_df_sample_sankey %>% group_by(PULocationID,DOLocationID) %>%
      summarize(Count = n())

final_table_ordered <- final_table[with(final_table, order(-Count)),]

final_table_head <- head(final_table_ordered, 50)

links <- data.frame(source = final_table_head$PULocationID,
                    target = final_table_head$DOLocationID,
                    value = final_table_head$Count) 

links <- links %>% mutate(source = recode(source, !!!names),
                          target = recode(target, !!!names))

links$source <- paste0(links$source,'_s')
links$target <- paste0(links$target,'_t')

nodes <- data.frame(name=c(as.character(links$source), as.character(links$target))) %>% 
  distinct()

links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1

links$group <- as.factor(links$source)
nodes$group <- as.factor(nodes$name)

p <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", iterations = 30,
              sinksRight=TRUE, LinkGroup="group", NodeGroup="group", units = 'pick-up', 
              fontSize = 12, fontFamily = "sans-serif") 

p <- htmlwidgets::prependContent(p, htmltools::tags$h3("The most common trips made by a yellow taxi"))

p
```


## Further analysis

Try to create a faceplotting of the PUL location of the three services as an example of further analysis.

```{r}
green_df <- read_csv("data/green_tripdata_2019-01.csv")
fhv_df <- read_csv("data/fhv_tripdata_2019-01.csv")
```

Edit the green cab dataframe in order to keep only the column usefull for the creation of the map.
```{r}
green_df_sample_taxy <- green_df[sample(nrow(green_df), 3219), ] %>%
  subset(select = c(PULocationID,DOLocationID)) 

green_nyc_PUL <- left_join(green_df_sample_taxy,map_nyc,"PULocationID") %>%
  subset(select = c(PULocationID,Shape_Leng,zone,borough, geometry)) %>%
  drop_na()
```

Edit the fhv dataframe by dropping all the na row and keep only the one usefull for the map.
```{r}
fhv_df <- drop_na(fhv_df)
fhv_sample <- fhv_df[sample(nrow(fhv_df), 3219), ]

fhv_nyc_PUL <- left_join(fhv_sample, map_nyc, "PULocationID") %>%
  subset(select = c(PULocationID,Shape_Leng,zone,borough, geometry)) %>%
  drop_na()
```

Plot the NYC map - green taxy PUL
```{r}
green_PUL_sample_taxy_map <- green_nyc_PUL %>% group_by(PULocationID) %>%
  summarize(count=n())

green_taxy_PUL_map <- left_join(green_PUL_sample_taxy_map,map_nyc, 'PULocationID') %>%
  subset(select = c(PULocationID, count, geometry))

ggplot() +
  geom_sf (data = map_nyc, size = 0.03) +
  geom_sf (data = green_taxy_PUL_map, mapping = aes(fill = count, geometry = geometry), size = 0.05) + 
  scale_fill_viridis_c()
```
